---
title: "Lierne Greening fall"
author: "Markus Fjellstad Israelsen"
date: "2023-12-08"
output: html_document
---


Load library
```{r Load Library}

library(countcolors)
library(camtrapR)
library(lubridate)
library(ggplot2)
library(dplyr)
library(writexl)
library(readxl)
library(gridExtra)
library(viridis)

```

Set up Green Fraction function for spring
```{r Green Fraction function, echo = TRUE, eval = TRUE}

GreenFraction = function(YearlyFolderDir){
  # Import exif tool
  exiftool_dir = "C:/Exiftool/"        
  exiftoolPath(exiftoolDir = exiftool_dir)
  
  # Define the directory for the image folders
  yearFoldersPath = YearlyFolderDir
  yearFoldersLength = list.files(yearFoldersPath)
  
  # For loop over the years
  for(i in 1:length(yearFoldersLength)){
    memCardFolders = list.files(paste0(yearFoldersPath, yearFoldersLength[i]))
    
    # For loop over the memory cards in each year
    for(j in 1:length(memCardFolders)){
      images = list.files(paste0(yearFoldersPath, yearFoldersLength[i], "/", memCardFolders[j]))
      images.db = which(images == "Thumbs.db")
      if(length(images.db) > 0){
        images = images[-images.db]
      }
      imagePath = paste0(yearFoldersPath, yearFoldersLength[i], "/", memCardFolders[j])
      MemCard = substring(imagePath, 76, nchar(imagePath))
      
      # For loop over each image, in each memory card, in each year
      imgDay = c()
      imgDayDF = data.frame("Date" = c(ymd("1970-01-01")), "Time" = "09:00:00", GreenFraction = c(NA), periodTempC = NA)
      tempDF = data.frame("Date" = c(ymd("1970-01-01")), "Time" = "09:00:00", tempC = NA)
      for (k in 1:length(images)){
        imgInfo = exifTagNames(inDir = imagePath, 1, images[k]) # extract image info
        imgDateTime = imgInfo[5, 3] # date and time
        Date = ymd(substring(imgDateTime, 1, 10))
        sTime = substring(imgDateTime, 20, 22) # correcting for summer time
        aTime = substring(imgDateTime, 12, 19)
        nTime = paste0(as.numeric(substring(aTime, 1, 2))+as.numeric(sTime), substring(aTime, 3, 8))
        Time = ifelse(nchar(nTime) == 7, paste0("0",as.numeric(substring(aTime, 1, 2))+as.numeric(sTime), substring(aTime, 3, 8)), nTime)
        tempC = as.numeric(substring(imgInfo[50, 3], 1, 2)) # Degrees centigrade
        tempDF[k, ] = c(as.character(Date), Time, tempC)
        
        # Filter out images in the time period of interest
        if(Time > "06:00:00" & Time < "18:00:00"){
          imgDay = append(imgDay, images[k])
          imgDayDF = rbind(imgDayDF, c(as.character(ymd(Date)), Time, NA, as.numeric(tempC)))
        }
      }
      imgDayDF = imgDayDF[-1, ]
      
      # For loop that extracts the fraction of green in each image
      for(h in 1:length(imgDay)){
    recImgGreenFraction = countcolors::rectangularRange(jpeg::readJPEG(paste0(imagePath, "/", imgDay[h])), upper = c(0.60, 1, 0.4), lower = c(0, 0.4, 0), color.pixels = FALSE, plotting = FALSE)
    recIGF = recImgGreenFraction$img.fraction * 100
    imgDayDF$GreenFraction[h] = recIGF
      }
      
      # Subset data frame based observations from just the whole hour (for instance, 12:00:00, 13:00:00 and so on.)
      imgDayDF = imgDayDF[grep(":00:00", imgDayDF$Time), ] 
      imgDayDF$GreenFraction = as.numeric(imgDayDF$GreenFraction)
      
      # Get the mean/median/min/max temperature for each image
      tempDF$tempC = as.numeric(tempDF$tempC)
      tempDF = tempDF[grep(":00:00", tempDF$Time), ]
      tempDfMean = tempDF %>% group_by(Date) %>% summarise(meanTempC = mean(tempC, na.rm = TRUE), minTempC = min(tempC, na.rm = TRUE), maxTempC = max(tempC, na.rm = TRUE), medianTempC = median(tempC, na.rm = TRUE))
      tempDfMean = tempDfMean %>% dplyr::select(Date, meanTempC, medianTempC, minTempC, maxTempC)
      tempDfMean$Date = as.Date(tempDfMean$Date)
      
      # Take the mean/median/min/max of the green fraction for each day
      imgDayDfMean = imgDayDF %>% group_by(Date) %>% summarise(meanGreenFraction = mean(GreenFraction, na.rm = TRUE), maxGreenFraction = max(GreenFraction, na.rm = TRUE), meanPeriodTempC = mean(as.numeric(periodTempC), na.rm = TRUE), medianPeriodTempC = median(as.numeric(periodTempC), na.rm = TRUE), minPeriodTempC = min(as.numeric(periodTempC), na.rm = TRUE), maxPeriodTempC = max(as.numeric(periodTempC), na.rm = TRUE))
      imgDayDfMean$Date = as.Date(imgDayDfMean$Date)
      imgDayDfMean = left_join(imgDayDfMean, tempDfMean, by = "Date")
      imgDayDfMean = imgDayDfMean %>% mutate(MemCard = MemCard)
      
      ## Export excel tables
      write_xlsx(imgDayDfMean, paste0("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/", "dailyGreen_", substring(yearFoldersLength[i], 14, 17), "_", substring(imagePath, 76, 89), ".xlsx"))
    }
  }
}

```

Set up Green Fraction function for LAST YEAR ONLY
```{r Green Fraction function, echo = TRUE, eval = TRUE}

GreenFractionLast = function(YearlyFolderDir){
  # Import exif tool
  exiftool_dir = "C:/Exiftool/"        
  exiftoolPath(exiftoolDir = exiftool_dir)
  
  # Define the directory for the image folders
  yearFoldersPath = YearlyFolderDir
  yearFoldersLength = list.files(yearFoldersPath)
  
  # For loop over the years
    memCardFolders = list.files(paste0(yearFoldersPath, max(yearFoldersLength)))
    
    # For loop over the memory cards in each year
    for(j in 1:length(memCardFolders)){
      images = list.files(paste0(yearFoldersPath, max(yearFoldersLength), "/", memCardFolders[j], "/"))
      images.db = which(images == "Thumbs.db")
      if(length(images.db) > 0){
        images = images[-images.db]
      }
      imagePath = paste0(yearFoldersPath, max(yearFoldersLength), "/", memCardFolders[j])
      MemCard = substring(imagePath, 76, nchar(imagePath))
      
      # For loop over each image, in each memory card, in each year
      imgDay = c()
      imgDayDF = data.frame("Date" = c(ymd("1970-01-01")), "Time" = "09:00:00", GreenFraction = c(NA), periodTempC = NA)
      tempDF = data.frame("Date" = c(ymd("1970-01-01")), "Time" = "09:00:00", tempC = NA)
      for (k in 1:length(images)){
        imgInfo = exifTagNames(inDir = imagePath, 1, images[k]) # extract image info
        imgDateTime = imgInfo[5, 3] # date and time
        Date = ymd(substring(imgDateTime, 1, 10))
        sTime = substring(imgDateTime, 20, 22) # correcting for summer time
        aTime = substring(imgDateTime, 12, 19)
        nTime = paste0(as.numeric(substring(aTime, 1, 2))+as.numeric(sTime), substring(aTime, 3, 8))
        Time = ifelse(nchar(nTime) == 7, paste0("0",as.numeric(substring(aTime, 1, 2))+as.numeric(sTime), substring(aTime, 3, 8)), nTime)
        tempC = as.numeric(substring(imgInfo[50, 3], 1, 2)) # Degrees centigrade
        tempDF[k, ] = c(as.character(Date), Time, tempC)
        
        # Filter out images in the time period of interest
        if(Time > "06:00:00" & Time < "18:00:00"){
          imgDay = append(imgDay, images[k])
          imgDayDF = rbind(imgDayDF, c(as.character(ymd(Date)), Time, NA, as.numeric(tempC)))
        }
      }
      imgDayDF = imgDayDF[-1, ]
      
      # For loop that extracts the fraction of green in each image
      for(h in 1:length(imgDay)){
    recImgGreenFraction = countcolors::rectangularRange(jpeg::readJPEG(paste0(imagePath, "/", imgDay[h])), upper = c(0.60, 1, 0.4), lower = c(0, 0.4, 0), color.pixels = FALSE, plotting = FALSE)
    recIGF = recImgGreenFraction$img.fraction * 100
    imgDayDF$GreenFraction[h] = recIGF
      }
      
      # Subset data frame based observations from just the whole hour (for instance, 12:00:00, 13:00:00 and so on.)
      imgDayDF = imgDayDF[grep(":00:00", imgDayDF$Time), ] 
      imgDayDF$GreenFraction = as.numeric(imgDayDF$GreenFraction)
      
      # Get the mean/median/min/max temperature for each image
      tempDF$tempC = as.numeric(tempDF$tempC)
      tempDF = tempDF[grep(":00:00", tempDF$Time), ]
      tempDfMean = tempDF %>% group_by(Date) %>% summarise(meanTempC = mean(tempC, na.rm = TRUE), minTempC = min(tempC, na.rm = TRUE), maxTempC = max(tempC, na.rm = TRUE), medianTempC = median(tempC, na.rm = TRUE))
      tempDfMean = tempDfMean %>% dplyr::select(Date, meanTempC, medianTempC, minTempC, maxTempC)
      tempDfMean$Date = as.Date(tempDfMean$Date)
      
      # Take the mean/median/min/max of the green fraction for each day
      imgDayDfMean = imgDayDF %>% group_by(Date) %>% summarise(meanGreenFraction = mean(GreenFraction, na.rm = TRUE), maxGreenFraction = max(GreenFraction, na.rm = TRUE), meanPeriodTempC = mean(as.numeric(periodTempC), na.rm = TRUE), medianPeriodTempC = median(as.numeric(periodTempC), na.rm = TRUE), minPeriodTempC = min(as.numeric(periodTempC), na.rm = TRUE), maxPeriodTempC = max(as.numeric(periodTempC), na.rm = TRUE))
      imgDayDfMean$Date = as.Date(imgDayDfMean$Date)
      imgDayDfMean = left_join(imgDayDfMean, tempDfMean, by = "Date")
      imgDayDfMean = imgDayDfMean %>% mutate(MemCard = MemCard)
      
      ## Export excel tables
      write_xlsx(imgDayDfMean, paste0("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/", "dailyGreen_", substring(max(yearFoldersLength), 14, 17), "_", substring(imagePath, 76, 89), ".xlsx"))
  }
}

```

Run the Greening Fraction Function for spring
```{r Run GF function, echo = TRUE, eval = FALSE}

# Supply the function with the directory path where the yearly greening images are stored (Kamerabilder 2019, Kamerabilder 2020..)
s = Sys.time()
GreenFraction(YearlyFolderDir = "P:/12179000_lirypetelemetri_i_lierne/ImagesGreenFraction/")
e = Sys.time()
runTime = e - s
runTime

# Start the greening function for the last year of camera images
s = Sys.time()
GreenFractionLast(YearlyFolderDir = "P:/12179000_lirypetelemetri_i_lierne/ImagesGreenFraction/")
e = Sys.time()
runTime = e - s
runTime

```

Import all the green fraction datasets and join them
```{r Import DF, echo = TRUE, eval = FALSE}

outputPath = "C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/"

greenFractionFiles = list.files(outputPath, pattern = "\\.xlsx$")
greenFractionList = lapply(paste0(outputPath, greenFractionFiles), read_excel)
greenFractionDF = bind_rows(greenFractionList) %>% mutate(YearDay = yday(Date), year = year(Date), monthNum = month(Date), month = month.abb[monthNum], day = day(Date))

# Remove all images after 25th of June
greenFractionDF = greenFractionDF %>% filter(YearDay <= 176)

write_xlsx(greenFractionDF, "C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/RmAfter25June/dailyGreenPixels.xlsx")

```

Check if it is best to use mean green fraction or max green fraction
```{r}

rl0019 = greenFractionDF %>% filter(year(Date) == 2021, MemCard == "RypeLierne0019")
ggplot(data = rl0019, aes(x = YearDay)) + geom_line(aes(y = meanGreenFraction), color = "magenta") + geom_line(aes(y = maxGreenFraction))

rl0017 = greenFractionDF %>% filter(year(Date) == 2021, MemCard == "RypeLierne0017")
ggplot(data = rl0017, aes(x = YearDay)) + geom_line(aes(y = meanGreenFraction), color = "magenta") + geom_line(aes(y = maxGreenFraction))

rl0013 = greenFractionDF %>% filter(year(Date) == 2021, MemCard == "RypeLierne0013")
ggplot(data = rl0013, aes(x = YearDay)) + geom_line(aes(y = meanGreenFraction), color = "magenta") + geom_line(aes(y = maxGreenFraction))

```
As expected the green fraction values are a bit higher with the "max green fraction", however, the start of the spring season and the trend/pattern of greening remains the same between the two methods. Using the mean green fraction from now on. 


Sampling quadrat 1 - Rl0009
```{r}

rl0009 = greenFractionDF %>% filter(MemCard == "RypeLierne0009")
rl0009.plot = ggplot(data = rl0009, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 1", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))
rl0009.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ1_2023.png", width=23, height=17, units="cm", dpi=600)

```

Sampling quadrat 2 - Rl0017
```{r}

rl0017 = greenFractionDF %>% filter(MemCard == "RypeLierne0017")
rl0017.plot = ggplot(data = rl0017, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 2", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))
rl0017.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ2_2023.png", width=23, height=17, units="cm", dpi=600)

```

Sampling quadrat 3 - Rl0013
```{r}
rl0013 = greenFractionDF %>% filter(MemCard == "RypeLierne0013")
rl0013.plot = ggplot(data = rl0013, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 3", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))
rl0013.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ3_2023.png", width=23, height=17, units="cm", dpi=600)

```

Sampling quadrat 4 - Rl0011
```{r}
rl0011 = greenFractionDF %>% filter(MemCard == "RypeLierne0011")
rl0011.plot = ggplot(data = rl0011, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 4", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))
rl0011.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ4_2023.png", width=23, height=17, units="cm", dpi=600)

```

Sampling quadrat 5 - Rl0015
```{r}
rl0015 = greenFractionDF %>% filter(MemCard == "RypeLierne0015")
rl0015.plot = ggplot(data = rl0015, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 5", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))
rl0015.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ5_2023.png", width=23, height=17, units="cm", dpi=600)

```

Sampling quadrat 6 - Rl0021
```{r}

rl0021 = greenFractionDF %>% filter(MemCard == "RypeLierne0021")
rl0021.plot = ggplot(data = rl0021, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 6", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))

rl0021.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ6_2023.png", width=23, height=17, units="cm", dpi=600)

```

Sampling quadrat 7 - Rl0019
```{r}

rl0019 = greenFractionDF %>% filter(MemCard == "RypeLierne0019")
rl0019.plot = ggplot(data = rl0019, aes(x = YearDay, y = meanGreenFraction, grouop = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per dag (%)") +
    labs(title = "Rute 7", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))
rl0019.plot
ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/SQ7_2023.png", width=23, height=17, units="cm", dpi=600)

```

Grid plot
```{r}

#grid.arrange(rl0009.plot, rl0017.plot, rl0013.plot, rl0011.plot, rl0015.plot, rl0021.plot, rl0019.plot,  nrow = 4, ncol = 2)

```


Mean of all sites 
```{r}

meanGreenFrac = greenFractionDF %>% group_by(YearDay, year) %>% summarise(meanSiteGreenFraction = mean(meanGreenFraction))
meanGreenFrac = arrange(meanGreenFrac, by= year)
meanGreenFrac$year = as.factor(meanGreenFrac$year)

ggplot(meanGreenFrac, aes(x = YearDay, y = meanSiteGreenFraction, group = year, color = year))+ geom_line()  + labs(color = "Year", x = "Year day", y = "Site mean 'green' fraction (%)") + geom_ribbon(aes(xmin = 60, xmax = 90, fill = "Mar", y = meanSiteGreenFraction), alpha = 0.10) + geom_ribbon(aes(xmin=90, xmax=120, fill = "Apr"), alpha = 0.10) + geom_ribbon(aes(xmin = 120, xmax = 151), alpha = 0.10, fill = "pink") + geom_ribbon(aes(xmin = 151, xmax = 181), alpha = 0.1)

#+ scale_color_manual(values=c("magenta", "steelblue", "dark green"))

ggplot(data = meanGreenFrac, aes(x = YearDay, y = meanSiteGreenFraction, group = as.factor(year), color = as.factor(year))) +
    geom_line(size = 1) + 
    xlab("Måned") +
    ylab("Gj.snittlig andel grønne piksler per år (%)") +
    labs(title = "Lifjellet", color = "År") +
    theme(
      legend.position = "top",
      title = element_text(size = 20),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 15),
      axis.title.x = element_text(size = 18),
      axis.title.y = element_text(size = 18),
      legend.text = element_text(size = 14),
      panel.background = element_rect(fill = "white", colour = "white"),
      panel.grid.major = element_line(size = 0.5, linetype = "solid", colour = "#e6e6e6")
    ) +
    scale_color_manual(values = c("#ffb25b", "#2dccd3", "#004f71", "#93328E", "#7A9A01")) + 
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(labels = c("Mar", "Apr", "Mai", "Jun", "Jul"))

ggsave("C:/Users/markus.israelsen/OneDrive - NINA/GitHub/Willow ptarmigan project/GreeningProject/LierneGreening/Greening output/Mean_SQ1-7_2023.png", width=23, height=17, units="cm", dpi=600)

```


